---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# riversCentralAsia

<!-- badges: start -->
<!-- badges: end -->

riversCentralAsia is an R Package with helper functions to load, manage and analyze hydrometeorological data from Central Asia. Currently, a relatively complete dataset of the Chirchik River Basin with decadal and monthly data on discharge, precipitation and temperature is included. The plan is to extend the data record to also include data from Chu and Talas rivers and from Amu Darya and Syr Darya, etc.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("hydrosolutions/riversCentralAsia")
```
## Example

This is a basic example which shows you how to visualize the included data

```{r example}
library(riversCentralAsia)
library(tidyverse)
library(timetk)
## basic example code
ChirchikRiverBasin # load data
ChirchikRiverBasin %>% group_by(code) %>% plot_time_series(date,data,
                                                           .interactive = FALSE,
                                                           .facet_ncol  = 2,
                                                           .smooth      = FALSE)
```

