---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# riversCentralAsia

## Summary

The R package riversCentralAsia includes a set of tools to facilitate and automate data preparation for hydrological modelling. It thus contributes to more reproducible modeling workflows and makes hydrological modeling more accessible to students. 

The package has been developed within the frame of a master level course on applied hydrological modelling in Central Asia and is extensively used in the open-source book [Modeling of Hydrological Systems in Semi-Arid Central Asia](https://hydrosolutions.github.io/caham_book/){target="_blank"} [@CAHAM:2022]. The workflows are further validated within the Horizon 2020 project [HYDRO4U](https://hydro4u.eu/){target="_blank"}. 

While the package has been developed for the Central Asia region, most of the functions are generic and can be used for modelling projects anywhere in the world. 

The most important functionalities of the package as well as the raw data that can be processed with the package are described in the the articles of the [project documentation site](https://hydrosolutions.github.io/riversCentralAsia/){target="_blank"} but the examples in the book [Modeling of Hydrological Systems in Semi-Arid Central Asia](https://hydrosolutions.github.io/caham_book/){target="_blank"} demonstrate the full range of functions available and how to use them in a workflow. 

## Statement of need

Data preparation comes before hydrological modelling and is actually one of the biggest work chunks in the modelling process. This package includes a number of helper functions that can be connected to efficient workflows that automatize the data preparation process for hydrological modelling. The functionality includes: 

- Efficient processing of present and future climate forcing, including hydro-meterological data from Central Asia ([time series and re-analysis data](https://github.com/hydrosolutions/riversCentralAsia/blob/master/.github/workflows/pkgdown.yaml){target="_blank"}) and down-scaling of ERA5 re-analysis data (a more advanced topic which is described in the [course book](https://hydrosolutions.github.io/caham_book/climate_data.html){target="_blank"})

- The [preparation of GIS layers for automated model generation](https://hydrosolutions.github.io/riversCentralAsia/articles/02-preparation-of-climate-forcing.html){target="_blank"} and chapter [Geospatial data](https://hydrosolutions.github.io/caham_book/geospatial_data.html){target="_blank"} in the course book

- [Volume area scaling of glaciers](https://hydrosolutions.github.io/riversCentralAsia/articles/04-glacier-functions.html){target="_blank"}

- [Post-processing of simulation results](https://hydrosolutions.github.io/riversCentralAsia/articles/03-RSMinerveIO.html){target="_blank"} 

- I/O interface with the hydrologic-hydraulic modelling software [RS MINERVE](https://crealp.ch/rs-minerve/){target="_blank"} which can be accessed within R using the package [RSMinerveR](https://github.com/hydrosolutions/RSMinerveR){target="_blank"}.  

While here, we focus on the description of the individual functions, the strengths of the package comes to play mostly when the functions are connected to automatize the data preparation process. These workflows are extensively documented in the book [Modeling of Hydrological Systems in Semi-Arid Central Asia](https://hydrosolutions.github.io/caham_book/){target="_blank"}. 


Currently, a relatively complete dataset of the Chirchik River Basin with decadal and monthly data on discharge, precipitation and temperature is included. 


## Related packages
The R package [RSMinverveR](https://github.com/hydrosolutions/RSMinerveR){target="_blank"} allows the running of the hydrologic-hydraulic modelling software [RS MINERVE](https://crealp.ch/rs-minerve/){target="_blank"} directly from R without using the RS MINERVE user interface. This functionality is for advanced R and RS MINERVE users that wish to further speed up their modelling workflow.     

## Installation
You can install the development version from [GitHub](https://github.com/hydrosolutions/riversCentralAsia/){target="_blank"} with:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_github("hydrosolutions/riversCentralAsia")
library(riversCentralAsia)
```

The package has many dependencies which will be installed alongside riversCentralAsia. To successfully install the package you need prior installations of the following packages: rlang, magrittr, stringr and purrr. Should the installation fail, test if you have the following dependencies installed.  
```{r, eval=FALSE}
find.package(c("rlang", "magrittr", "stringr", "purrr"))
```

If they are not yet available on your system, please install them using the following commands. 
```{r, eval=FALSE}
install.packages(c("rlang", "magrittr", "stringr", "purrr"))
```
All other dependencies are installed automatically. Some users reported, that they had to install the package exactextractr manually. 

Note that windows users require a working installation of [RTools](https://cran.r-project.org/bin/windows/Rtools/){target="_blank"} to install packages from github. 

We recommend testing of the riversCentralAsia package before using it to make sure all dependencies integrate with the riversCentralAsia package as they should. The following step-by-step instructions explain how to do this:   

1. Go to the [packages github repository](https://github.com/hydrosolutions/riversCentralAsia), click the green code button and select download zip. This will download a zip folder with the package code to your Download folder.    
2. Extract the package files to a location of your preference. More detailed instructions for this step: right-click on the zip folder and select "extract all"in the pop-up window. Select a location to extract the files to and wait for the process to finish.   
3. Navigate into the riversCentralAsia folder to in which the freshly downloaded package data is stored and open the riversCentralAsia R project file. This will open the project in your R GUI.   
4. In the R console, type ```devtools::test()``` and hit enter. The tests will run for about a minute. If the test return ```[ FAIL 0 | WARN 0 | SKIP 0 | PASS 86 ]``` at the end, all test were successful and the package will run without problems on your system. Should one of the tests fail, please make sure all the package dependencies are up to date. If the problem persists, please file an [issue](https://github.com/hydrosolutions/riversCentralAsia/issues) including the error message you get so we can look at the problem and hopefully solve it.    
5. If you don't need the source codes any more, you can safely delete the zip file and the folder with the extracted riversCentralAsia files. 


## Community guidelines
Please consult the documentation and the examples provided in the [package documentation](https://hydrosolutions.github.io/riversCentralAsia/index.html){target="_blank"} and in the open-source course book [Modeling of Hydrological Systems in Semi-Arid Central Asia](https://hydrosolutions.github.io/caham_book/){target="_blank"}. 

For problems using the functions of for suggestions, please use the [issue tool](https://github.com/hydrosolutions/riversCentralAsia/issues){target="_blank"}. 

## How to cite
Please cite the package as: 

Tobias Siegfried, & Beatrice Marti (2021): riversCentralAsia <version number>. https://doi.org/10.5281/zenodo.4667422 

[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.4667422.svg)](https://doi.org/10.5281/zenodo.4667422)



## Examples

This is a basic example which shows you how to visualize some of the included data.


```{r example, echo=TRUE, eval=FALSE}
library(riversCentralAsia)
library(tidyverse)
library(timetk)

# Loading and visualising discharge data
ChirchikRiverBasin  # load data
ChirchikRiverBasin |> 
  # Filter for the data type, here discharge "Q"
  dplyr::filter(type == "Q") |> 
  drop_na() |> 
  group_by(river) |> 
  plot_time_series(
    date,
    data,
    .interactive = FALSE,
    .facet_ncol = 2,
    .smooth = FALSE, 
    .y_lab = "Discharge [m3/s]", 
    .x_lab = "Year", 
    .title = "Discharge time series in the ChirchikRiverBasin data set"
    )
```

## Mentions
The package is used extensively in the course book [Modeling of Hydrological Systems in Semi-Arid Central Asia](https://hydrosolutions.github.io/caham_book/){target="_blank"} (Siegfried & Marti, 2022). 

The workflows presented in the course book, using the riversCentralAsia package, are further validated within the Horizon 2020 project [HYDRO4U](https://hydro4u.eu/){target="_blank"} where future small hydro power potential is evaluated using hydrological modelling. 

For R & RS MINERVE users, the package [RSMinverveR](https://github.com/hydrosolutions/RSMinerveR){target="_blank"} is recommended which allows the interfacing between R and RS MINERVE (with examples based on the Visual Basic Script examples by CREALP).  

## Acknowledgement
The preparation of the course book and thus the preparation of the package was financially supported by the Swiss Agency for Development and Cooperation, the German Kazakh University in Almaty and hydrosolutions. 

This R package was submitted to the Journal of Open Source Software. We thank @tonyewong and @mengqi-z for their valuable inputs during the review of the package and @crvernon for their editing work. 

## References
Tobias Siegfried & Beatrice Marti (2022): Modeling of Hydrological Systems in Semi-Arid Central Asia (\url{https://hydrosolutions.github.io/caham_book/}). DOI: 10.5281/zenodo.6350042
