---
title: "Introduction to the package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to the package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(
  rmarkdown.html_vignette.check_title = FALSE
)
```

This package aims at streamlining pre-and post-processing of various types of data for hydrological modelling in Central Asia. It is under continuous development. 

Several use-cases for the packages functionality are demonstrated in the CAHAM book [Modeling of Hydrological Systems in Semi-Arid Central Asia](https://hydrosolutions.github.io/caham_book/){target="_blank"}. 

## Installation
```{r, message=FALSE, warning=FALSE, error=FALSE}
# install.packages("devtools")
devtools::install_github("hydrosolutions/riversCentralAsia", 
                         quiet = TRUE)
library(riversCentralAsia, quietly = TRUE)
```

## Data
The package includes an example data set with discharge time series as well as temperature and precipitation time series from the runoff formation zone of the Chirchiq river basin. 
```{r, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
library(riversCentralAsia, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(timetk, quietly = TRUE)
ChirchikRiverBasin  # load data
ChirchikRiverBasin |> 
  filter(type == "Q", 
         code %in% c("16279", "16290", "16298", "16300")) |> 
  group_by(code, station, river) |> 
  drop_na() |> 
  plot_time_series(
    date, data,
    .interactive = TRUE,
    .facet_ncol  = 2, 
    .facet_collapse = TRUE, .facet_collapse_sep = " - ", 
    .smooth      = FALSE, 
    .y_lab = "Q [m3/s]")
```

## Tools
The package `riversCentralAsia` includes a variety of functions to facilitate data pre- and post-processing for hydrological and hydraulic modelling with RS Minerve in Central Asia.    
  
[RS Minerve](https://crealp.ch/rs-minerve/){target="_blank"} is a free hydrological and hydraulic modelling software developed in Switzerland by CREALP and partners. 

In summary, the functions can be grouped into:   
* Reading and writing of input and output files of RS Minerve    
* Pre- and post-processing of data for hydrolgical modelling with RS Minerve  
* Glacier modelling tools  

##









