---
title: "I/O for RS Minerve"
output: rmarkdown::html_vignette
bibliography: [references.bib, HMA_bibliography.bib]
vignette: >
  %\VignetteIndexEntry{I/O for RS Minerve}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_knit$set(
  rmarkdown.html_vignette.check_title = FALSE
)
```

[RS Minerve](https://crealp.ch/rs-minerve/){target="_blank"} is a free to use hydrologic-hydraulic modelling software. It's accessibility makes is especially suitable for typical Hydromet tasks like flow forecasting. The package riversCentralAsia includes a number of functions to read and write files which interact with RS Minerve. 
These tools are interesting for automated writing of forcing files and for reading post-processing and visualization of simulation results or model parameters. 

Using the package `RSMinerveR`, the software can be run from inside R. 

For details on model components please refer to the [RS Minerve manuals](https://crealp.github.io/rsminerve-releases/){target="_blank"}. 

```{r fig.cap="Storage coefficient of the upper reservoirs in the HBV models by zone."}
library(tidyverse)
library(riversCentralAsia)

parameters_orig <- readRSMParameters("../../atbashy_glacier_demo_data/RSMINERVE/Atbaschy_PAR_postCal.txt") |> 
  dplyr::filter(Parameters == "Ku [1/d]") 

ggplot(parameters_orig) + 
  geom_col(aes(Zone, Values), position = "dodge") + 
  ylab(parameters_orig$Parameters[1]) + 
  theme_bw()
```

The following shows an example how the model parameters can be changed and written to a RS Minerve-readable format. 

```{r}
# Doubble all Ku-values in zone A: 
parameters <- readRSMParameters("../../atbashy_glacier_demo_data/RSMINERVE/Atbaschy_PAR_postCal.txt") |> 
  mutate(Values = ifelse(Parameters == "Ku [1/d]" & Zone == "A", Values * 2, 
                         Values))

comparison <- parameters_orig |> 
  mutate(Set = "Original") |> 
  add_row(parameters |> dplyr::filter(Parameters == "Ku [1/d]") |> 
            mutate(Set = "Modified"))

ggplot(comparison) + 
  geom_col(aes(Zone, Values, fill = Set), position = "dodge") + 
  scale_fill_viridis_d() + 
  ylab(parameters_orig$Parameters[1]) + 
  theme_bw()

# writeRSMParameters(parameters, "Parameters_modified_Ku.txt")
```

Another useful application for the RS Minerve I/O functions is the reading of simulation results and the analysis of the partitioning of the total basin discharge into the individual fluxes. The code below illustrates how this may be done: 

Writing a selection node file that can be imported to RS Minerve where the selection can be used to export the selected simulation results. This is useful for large models with many model objects where it is cumbersome to manually toggle the desired results in RS Minerve.  

```{r, eval=FALSE}
# Write check node file
hbv_var_list <- c("Qr (m3/s)", "Qu (m3/s)", "Ql (m3/s)", "SWE (m)", "HUM (m)", 
                  "SU (m)", "SL (m)", "Peq (m/s)")
station_var_list <- c("P (m/s)")

data <- tibble(
  Model = rep("Model Atbaschy", 
              length(Object_IDs$name)*length(hbv_var_list) + 
                length(Object_IDs$name)*length(station_var_list)), 
  Object = c(rep("HBV92", (length(Object_IDs$name)*length(hbv_var_list))),
             rep("Station", (length(Object_IDs$name)*length(station_var_list)))), 
  ID = c(rep(Object_IDs$name, each = length(hbv_var_list)), 
         paste("Station", rep(Object_IDs$name, each = length(station_var_list)), 
               sep = " ")), 
  Variable = c(rep(hbv_var_list, length(Object_IDs$name)), 
               rep(station_var_list, length(Object_IDs$name)))
)

writeSelectionCHK(filepath = paste0(data_path, 
                                    "RSMINERVE/FluxesAndStates.chk"), 
                  data = data, 
                  name = "Fluxes and states")
```

We then read the simulation results. 

```{r, message=FALSE, warning=FALSE, error=FALSE, fig.cap="Average seasonal solid and liquid precipitation and contribution of the HBV model compartments and glacier melt to river discharge."}

data_path <- "../../atbashy_glacier_demo_data/"

# Fluxes and model states per elevation bands
fs <- readResultCSV(paste0(data_path, 
                          "RSMINERVE/Atbaschy_Results_Fluxes_and_States.csv")) |> 
  mutate(model = gsub("Station ", "", model))

Tth <- 1  # HBV model parameter for partitioning of solid and liquid precipitation
forcing <- readForcingCSV(paste0(data_path, 
                                 "RSMINERVE/hist_obs_rsm.csv")) |> 
  dplyr::select(-Category) |> 
  pivot_wider(names_from = c(Sensor, Unit), names_sep = "_", values_from = Value)  |> 
  mutate(Month = month(Date), 
         P_mmd = `P_mm/d`, 
         Rain_mmd = ifelse(T_C >=  Tth, `P_mm/d`, 0), 
         Snow_mmd = ifelse(T_C < Tth, `P_mm/d`, 0)) |> 
  # Long-term mean per HRU  
  group_by(Month, Station) |> 
  summarise(P_mmd = mean(P_mmd, na.rm = TRUE), 
            Rain_mmd = mean(Rain_mmd, na.rm = TRUE), 
            Snow_mmd = mean(Snow_mmd, na.rm = TRUE)) |> 
  # Sum per Basin
  group_by(Month) |> 
  summarise(P_mmd = sum(P_mmd, na.rm = TRUE), 
            Rain_mmd = sum(Rain_mmd, na.rm = TRUE), 
            Snow_mmd = sum(Snow_mmd, na.rm = TRUE)) |> 
  ungroup() |> 
  dplyr::select(-P_mmd) |> 
  pivot_longer(-Month, names_to = "Precipitation", values_to = "Values") |> 
  mutate(Precipitation = ifelse(Precipitation == "Rain_mmd", "Liquid", "Solid")) |> 
  rename(Component = Precipitation)

forcing_plot <- ggplot(forcing) + 
  geom_col(aes(Month, Values, fill = Component), position = "stack") + 
  ylab("Precipitation [mm/d]") + 
  scale_x_continuous(breaks = c(1:12), 
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", 
                                "O", "N", "D")) + 
  theme_bw()


# Glacier discharge is already aggregated to sub-basin level 
gl <- readResultCSV(paste0(data_path, 
                           "RSMINERVE/Atbaschy_Results_Glacier_Sources.csv")) |>
  mutate(variable = "Qgl", 
         model = gsub("glaciers_", "", model), 
         model = paste(toupper(substr(model, 1, 1)), 
                       substr(model, 2, nchar(model)), sep=""), 
         Month = month(date)) |> 
  # Calculate mean flux per HRU
  group_by(Month, variable, model) |> 
  summarise(value = mean(value), na.rm = TRUE) |>
  # Sum over basin
  group_by(Month, variable) |> 
  summarise(value = sum(value, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(variable = factor(variable, 
                           levels = c("Qgl", "Qr", "Qu", "Ql")))

mean_monthly_fluxes <- fs |> 
  dplyr::filter(unit == "m3/s", 
                date > as_date("1990-01-01")) |> 
  mutate(Month = month(date)) |> 
  group_by(Month, variable, model) |> 
  summarise(value = mean(value), na.rm = TRUE) |> 
  group_by(Month, variable) |> 
  summarise(value = sum(value, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(variable = factor(variable, 
                           levels = c("Qgl", "Qr", "Qu", "Ql"))) |> 
  add_row(gl)

discharge_plot <- ggplot(mean_monthly_fluxes) + 
  geom_area(aes(Month, value, fill = variable), position = "stack") + 
  scale_fill_viridis_d(name = "Component", direction = -1) +
  ylab("Discharge [m3/s]") + 
  scale_x_continuous(breaks = c(1:12), 
                     labels = c("J", "F", "M", "A", "M", "J", "J", "A", "S", 
                                "O", "N", "D")) + 
  theme_bw()
  
gridExtra::grid.arrange(forcing_plot, discharge_plot, ncol = 1)
```

And last but not least a modeler might wish to produce flow duration curves based on simulation results. Flow duration curves are used, for example, for the design of hydraulic infrastructure.

```{r, message=FALSE, warning=FALSE, error=FALSE, fig.cap="Flow duration curve of simulated monthly discharge at the planning location for a small hydropower plant. Black: Average flow duration curve over the period 1990-2021, grey: flow duration curve for the years 1990 - 2011."}
# Calculate the flow duration curve
Qtot <- fs |> 
  dplyr::filter(unit == "m3/s", 
                date > as_date("1990-01-01")) |> 
  group_by(date) |> 
  summarise(Qtot_m3s = sum(value)) |> 
  ungroup()

fdc <- computeAnnualFlowDurationCurve(data = Qtot, column = "Qtot_m3s", 
                                      date = "date") |> 
  dplyr::filter(Ma <= 12) |> 
  mutate(Ma = as.numeric(Ma), 
         Year = factor(Year))

fdc_mean <- fdc |> 
  group_by(Ma) |> 
  summarise(Qtot_m3s = mean(Qtot_m3s)) |> 
  ungroup()
  
ggplot(fdc) + 
  geom_line(aes(Ma, Qtot_m3s, colour = Year), size = 0.6, alpha = 0.2,
            show.legend = FALSE) + 
  geom_line(data = fdc_mean, 
            aes(Ma, Qtot_m3s)) + 
  ylab("Discharge [m3/s]") + 
  xlab("Duration [months]") + 
  scale_colour_manual(values = rep("gray", length(unique(fdc$Year)))) + 
  scale_x_continuous(breaks = c(1:12), 
                     labels = c(1:12)) + 
  theme_bw()
```

In gray we see the flow duration curves of the individual years and in black the mean flow duration curve.


