---
title: "Glaciers 02 - Temperature index model"
editor: visual
bibliography: references.bib
output: rmarkdown::html_vignette
fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Glaciers 02 - Temperature index model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, message=FALSE, warning=FALSE}
library(riversCentralAsia)
library(tidyverse)
library(lubridate)
library(sf)
library(raster)
library(exactextractr)
```

# Temperature index model

@hock_temperature_2003 describes several variants of the temperature index model for simulating glacier melt. The riversCentralAsia Package implements the temperature index melt model described in [@hock_temperature_2003] in the function \code{glacierMelt_TI}.

<!-- # TODO: I would you different notation here for the emlt factor. In the published version it is confusing as MF looks like M F where M is the M on the f_{m} -->

$$
M = \biggl\{ \begin{array}{l, l}
0, & T < T_{threshold} \\
MF \cdot \left( T - T_{threshold} \right), & T >= T_{threshold}
\end{array} 
$$

where $M$ is the glacier melt in $mm/d$, $T$ is the daily average temperature in $^{\circ} C$. The two parameters $MF$ and $T_{\text{threshold}}$ refer to the melt factor and the threshold temperature above which glacier melt occurs and need to be calibrated. They have the units $\frac{mm}{^{\circ} C \cdot d}$ and $^{\circ} C$ respectively. Glacier melt is calculated in daily time steps.

# Glacier mass balance

We simplify the glacier mass balance to the following equation

$$
\Delta S = P-M = A_{imbal}
$$

where the change of water storage ($\Delta S$) is equal to the precipitation ($P$, assumed solid) minus the glacier melt ($M$). Typically melt excesses precipitation and we have negative imbalance ablation, indicating glacier storage loss. The glacier mass balance is calculated in annual time steps. We thereby refer to the hydrological year starting on October 1st of the previous year to take a full accumulation and ablation season into account.

# Glacier volume development

As glaciers melt, their volume changes. This has to be taken into account for the long-term simulation of glacier discharge. The glacier volume develops according to the mass change. To determine the initial glacier volume we use the area of the geometry of the [RGI v6.0 data set](glaciers-01-intro.html) and multiply it with the average [thickness of the glacier](glaciers-01-intro.html). Glaciers larger than 1km2 we sub-divide into elevation bands of 100m altitude to account for elevation dependent temperature forcing.

For annual time step $t$, the evolution of the glacier volume is calculated as follows:

$$
A(t) = \text{glacierArea_RGIF}\bigl(V(t)\bigr)
$$

$$
Q_{glacier}(t) = M(t) \cdot A(t)
$$

$$
V(t+1) = V(t) + \Delta S = V(t) + \text{glacierImbalAbl}\bigl(Q_{\text{glacier}}(t)\bigr) 
$$

Note that the imbalance ablation (or $\Delta S$) is negative for glacier mass loss and positive for glacier accumulation.

$Q_{\text{glacier}}$ can be calibrated against glacier discharge derived from the Miles & Hugonnet data sets.

The function \code{glacierArea_RGIF()} is an empirical scaling function analogue the inverse of the scaling function derived by @erasov_1968 but based on the modern RGI v6.0 glacier geometries and the glacier thickness data set by @farinotti_consensus_2019. The package \code{riversCentralAsia} implements volume-area and area-volume scaling functions based on both, Erashov and RGI-Farinotti data, allowing the estimation of glacier areas based on glacier volumes (\\code{glacierArea_Erasov} and \\code{glacierArea_RGIF}) and estimations of glacier volumes based on glacier areas (\\code{glacierVolume_Erasov} and \\code{glacierVolume_RGIF}).

The function \code{glacierImbalAbl} is an empirical scaling function relating total glacier melt to imbalance ablation. It is derived from the glacier discharge data set by @miles_health_2021.

# Propagation of uncertainty

The initial glacier volume of each glacier is attributed an uncertainty of p/m 26%. This number is based on the average uncertainty of the glacier volume per RGI region reported in @farinotti_consensus_2019 (Table 1). The error is the same for RGI regions 13 and 14. The uncertainty of the area of the RGI v6.0 glacier outlines is not known. We therefore assume that the error of the glacier volume stems to 50% from the estimation of the glacier thickness and to 50% from the glacier area. We further assume that the errors of the glacier area and glacier thickness are un-correlated and can thus estimate the uncertainties of the glacier area data and the glacier thickness data to be p/m 13% each.

For the non-linear relationships \code{glacierVolume_RGIF} and \code{glacierArea_RGIF}, the standard deviation of the residuals of the fit was computed. The estimated error of the fit is assumed to be equal plus/minus twice the standard deviation of the residuals and yields 31% and 53% respectively. The residuals are not normally distributed, their actual distribution in unknown. We assume uncorrelated errors and therefore add the error of the function input to the error of the fit to obtain the error of the function output.

$$
\varepsilon_{V} = \varepsilon_{A} + \varepsilon_{\text{glacierVolume_RGIF}} = 0.26\% + 0.31\% = 0.57\%
$$

$$
\varepsilon_{A} = \varepsilon_{V} + \varepsilon_{\text{glacierArea_RGIF}} = 0.26\% + 0.53\% = 0.79\%
$$

We do not have an estimate for the error of the temperature index model. We assume a relative error of 2, indicating that the estimated glacier melt is within a range of plus/minus 2 times it's value.

The glacier discharge relation likewise have estimation errors. For glacier growth we have to assume an average relative error of 1.42% and for glacier melt an average relative error of 0.63%.

$$
\varepsilon_{Q_{\text{gl,growth}}} = \varepsilon_{dhdt} + \varepsilon_{\text{glacierDischarge_HM,growth}} = 0.74\% + 1.42\% = 2.16\%
$$

$$
\varepsilon_{Q_{\text{gl,melt}}} = \varepsilon_{dhdt} + \varepsilon_{\text{glacierDischarge_HM,melt}} = 0.74\% + 0.63\% = 1.37\%
$$

Similarly the error of the imbalance ablation amounts to 1.27% yielding an error of the imbalance ablation of:

$$
\varepsilon_{Q_{\text{imb,melt}}} = \varepsilon_{Q_{\text{gl,melt}}} + \varepsilon_{\text{glacierImbalAbl}} = 1.37\% + 1.27\% = 2.64\%
$$

The errors stated above relate to initial estimates of the glacier areas, volumes, total ablation and imbalance ablation. The uncertainties are not propagated through each time step of the model as the errors cannot be assumed to be uncorrelated.

```{r errstats}
# Relative error estimates for the initial state in %
error_stats = tibble(
  sV = 0.26,  # Glacier volume
  sA = 0.13,  # Glacier area
  sTh = 0.13,  # Glacier thickness
  sMelt = 2,  # Glacier melt
  sVolumeRGIF = 0.31,  # Calculating glacier volume from glacier area
  sAreaRGIF = 0.53,  # Calculating glacier area from glacier volume
  sQglgrowth = 2.16,  # Calculating total glacier discharge from glacier thinning rates for growing glaciers
  sQglmelt = 1.37,  # Calculating total glacier discharge from glacier thinning rates for melting glaciers
  sImbal = 1.27)  # Calculating glacier imbalance ablation from total glacier discharge
```

# Forcing

There will be a separate Vignette to demonstrate how to prepare the forcing. For now, we provide you with the pre-processed forcing data and give you just a brief overview over the data source.

## Historical temperatures

As meteorological data for high elevations in Central Asia is very scarce we use the CHELSA v2.1 data set. This is a global data set of forcing data for hydrolgical models based on ERA5 but corrected for biases and for high elevations. The daily CHELSA forcing has been cut to the Central Region by the originator of the data set, D. Karger, WSL.

```{r histobs, message=FALSE, warning=FALSE}
#| fig.cap = "Daily temperature time series of a glacier. Data source: CHELSA." 
#| out.width = 90%
hist_obs <- readRDS(file = "hist_obs_glacier_tas.rds")

# Plot the temperature time series for a given glacier/elevation band
glacier <- "RGI60-13.08930_1"
ggplot(hist_obs) + 
  geom_line(aes(date, get(glacier))) + 
  labs(x = "Date", y = "T [deg C]", title = glacier) +
  theme_bw()
```

## Future temperatures

Future temperature development per glacier or elevation band is extracted from the 4 CMIP6 GCM models with highest priorities for the region () downloaded from COPERNICUS. We take 4 socioeconomic scenarios into account, covering 4 different emission scenarios. The temperatures of the climate models are bias corrected using the CHELSA data and a quantile mapping method.

```{r futsim, message=FALSE, warning=FALSE}
#| fig.cap = "Annual historical (CHELSA) and future (CMIP6, Copernicus) temperature for a glacier. " 
#| out.width = 90%
fut_sim <- readRDS(file = "fut_sim_glacier_tas_qmapped.RDS")

# Plot the temperature time series for a given glacier/elevation band
glacier <- "RGI60-13.08930_1"

# Extract the temperature for the selected glaciers for all GCMs and SSPs
scenarios <- names(fut_sim)
fut_temp <- NULL
for (scenario in scenarios) {
  fut_temp <- rbind(fut_temp, 
                    fut_sim[[scenario]] |> 
                      dplyr::select(Date, all_of(glacier)) |> 
                      mutate(Scenario = scenario))
}
fut_temp <- fut_temp |> 
  mutate(Hyear = hyear(Date)) |> 
  group_by(Hyear, Scenario) |> 
  summarise(Date = first(Date), 
            Temp = mean(get(glacier))) |> 
  separate(Scenario, into = c("GCM", "SSP"), sep = "_") |> 
  ungroup() |>
  dplyr::filter(Hyear > min(Hyear) & Hyear < max(Hyear), 
                GCM != "IPSL-CM6A-LR")

# Plot annual data for better readability
ggplot() + 
  geom_line(data = hist_obs |> 
              mutate(Hyear = hyear(date)) |> 
              group_by(Hyear) |> 
              summarise(date = first(date), 
                        Temp = mean(get(glacier))) |> 
              ungroup() |> 
              dplyr::filter(Hyear > min(Hyear) & Hyear < max(Hyear)), 
            aes(date, Temp)) + 
  geom_line(data = fut_temp, aes(Date, Temp, colour = GCM, 
            linetype = SSP)) + 
  scale_colour_viridis_d() + 
  labs(x = "Date", y = "T [deg C]", title = glacier) +
  theme_bw()
```

# Demonstration

We demonstrate the above described method with the data from the Atbashy basin.

## Calculate daily glacier melt

### Glacier geometry

```{r geometry, message=FALSE, warning=FALSE}
# Load a shape file with elevation bands for the larger glaciers.
data <- vignette_glacier_melt

# Only keep the variables we need for this analysis
rgi_elbands <- data$rgi_elbands |> 
  dplyr::select(RGIId, Area, elvtn_b) |> 
  rename(Area_tot_glacier_km2 = Area) |> 
  mutate(ID = paste0(RGIId, "_", elvtn_b))

# Get mean elevation of each glacier/elevation band from DEM
rgi_elbands$z_masl <- exact_extract(data$dem, rgi_elbands, "mean", progress = FALSE)

# Update the glacier area within the basin boundaries
rgi_elbands$A_km2 <- as.numeric(st_area(rgi_elbands))*10^(-6)
glaciers <- unique(rgi_elbands$RGIId)

# Glacier thickness for elevation bands
rgi_elbands$thickness_m = exact_extract(data$glacier_thickness, 
                                        rgi_elbands, "mean", progress = FALSE)
```

### Forcing

When simulating glacier melt, you would typically extract a temperature time series for each glacier/elevation band. In this vignette, we only have 1 time series $T_0$ at mean glacier elevation $z_0$ to save storage space. We therefore first have to generate synthetic temperature time series for each glaciers/elevation band $i$ using a typical temperature lapse rate $r_{\text{lapse}}$ for mountainous regions as follows:

$$
\frac{\Delta T}{\Delta z} = \frac{T_0-T_i}{z_0-z_i} = r_{\text{lapse}} = -\frac{6.5 ^{\circ}C}{1000 m}
$$

$$
T_{i} = T_{0} + r_{\text{lapse}} \cdot (z_{0} - z_{i})
$$

```{r temperature_elevation_bands, message=FALSE, warning=FALSE}
#| fig.cap = "Synthetic forcing for the elevation bands of a glacier. " 
#| out.width = 90%
glacier <- "RGI60-13.08930_1"
z_glacier <- rgi_elbands$z_masl[rgi_elbands$ID == glacier]
temperature_laps_rate <- -6.5/1000  # deg K / 1000 m 

# Generate synthetic temperature time series for each glacier & elevation band in 
# the basin. 
z_glaciers <- rgi_elbands |> 
  st_drop_geometry() |> 
  dplyr::select(ID, z_masl) |> 
  pivot_wider(names_from = ID, values_from = z_masl)

hrus <- rgi_elbands$ID
t_glaciers <- as.matrix(hist_obs)
colnames(t_glaciers) <- NULL
for (hru in c(2:length(hrus))) {
  t_glaciers <- cbind(t_glaciers, as.matrix(as.numeric(t_glaciers[, 2])) + 
    temperature_laps_rate * 
    (as.numeric(z_glaciers[hru]) - as.numeric(z_glaciers[1])))
} 
colnames(t_glaciers) <- c("date", hrus) 
t_glaciers <- as_tibble(t_glaciers)
t_glaciers$date <- as_date(t_glaciers$date)
t_glaciers <- t_glaciers |> 
  mutate(across(starts_with("RGI"), ~as.numeric(.x)))

ggplot(t_glaciers[, 1:10] |> 
         pivot_longer(-date, names_to = "ID", values_to = "Temp")) + 
  geom_line(aes(date, Temp, colour = ID)) + 
  scale_colour_viridis_d() + 
  labs(x = "Date", y = "T [deg C]") + 
  theme_bw()
```

### Calculating glacier melt

```{r glaciermeltcalculation, message=FALSE, warning=FALSE}
#| fig.cap = "Daily glacier melt per elevation band." 
#| out.width = 90%
MF_small = 4
MF_large = 0.5
threshold_temperature = 0
Area <- rgi_elbands |> 
  st_drop_geometry() |> 
  dplyr::select(ID, A_km2) |> 
  pivot_wider(names_from = ID, values_from = A_km2)
# Assign different melt factors to large and small glaciers. 
MF <- Area |> 
  mutate(across(everything(), ~MF_large), 
         across(ends_with("_1"), ~MF_small))

melt <- glacierMelt_TI(temperature = t_glaciers |> dplyr::select(-date),
                       MF = MF,
                       threshold_temperature = threshold_temperature)
melt <- as_tibble(melt) |> 
  mutate(date = t_glaciers$date) |> 
  relocate(date, .before =  where(is.numeric))

ggplot(melt[, 1:10] |> 
         pivot_longer(-date, names_to = "ID", values_to = "Melt")) + 
  geom_line(aes(date, Melt, colour = ID)) + 
  scale_colour_viridis_d() + 
  labs(x = "Date", y = "Melt [mm/d]") + 
  theme_bw()
```

### Compare to measured glacier melt

Aggregate the daily glacier melt to annual data.

```{r glaciermeltcomparison, message=FALSE, warning=FALSE}
#| fig.cap = "Daily glacier melt per elevation band." 
#| out.width = 90%
melt_a_eb <- melt |> 
  pivot_longer(-date, names_to = "ID", values_to = "Melt") |> 
  mutate(Hyear = hyear(date)) |> 
  group_by(Hyear, ID) |> 
  summarise(date = first(date), 
            Melt = sum(Melt), 
            .lb_Melt = ifelse(Melt*(1-error_stats$sMelt)<0, 0, 
                              Melt*(1-error_stats$sMelt)), 
            .ub_Melt = Melt*(1+error_stats$sMelt))|> 
  ungroup() 
melt_a <- melt_a_eb |> 
  separate(ID, into = c("RGIId", "elB"), sep = "_") |> 
  group_by(Hyear, RGIId) |> 
  summarise(date = first(date), 
            Melt = sum(Melt), 
            .lb_Melt = ifelse(Melt*(1-error_stats$sMelt)<0, 0, 
                              Melt*(1-error_stats$sMelt)), 
            .ub_Melt = Melt*(1+error_stats$sMelt)) |> 
  ungroup() |> 
  dplyr::filter(Hyear > min(Hyear) & Hyear < max(Hyear))

glaciers_hugonnet <- data$glaciers_hugonnet |> 
  mutate(Qgl_m3a = glacierDischarge_HM(dhdt), 
         .lb_Qgl_m3a = ifelse(
           dhdt > 0, 
           ifelse(Qgl_m3a*(1-error_stats$sQglgrowth)<0, 0, 
                  Qgl_m3a*(1-error_stats$sQglgrowth)), 
           ifelse(Qgl_m3a*(1-error_stats$sQglmelt)<0, 0, 
                  Qgl_m3a*(1-error_stats$sQglmelt))),
         .ub_Qgl_m3a = ifelse(dhdt > 0, 
                              Qgl_m3a*(1+error_stats$sQglgrowth), 
                              Qgl_m3a*(1+error_stats$sQglmelt)))
melt_obs_a <- glaciers_hugonnet |> 
  dplyr::filter(RGIId %in% glaciers[6:9], 
                period == 1)
  
ggplot() + 
  geom_ribbon(data = melt_a |> 
                dplyr::filter(RGIId %in% glaciers[6:9]), 
              aes(date, Melt/1000, ymin = .lb_Melt/1000, ymax = .ub_Melt/1000, 
                  fill = RGIId), colour = NA, alpha = 0.2) + 
  geom_ribbon(data = melt_obs_a |> 
                dplyr::filter(RGIId %in% glaciers[6:9]), 
              aes(start, -dmdtda, 
                  ymin = -dmdtda-err_dmdtda, 
                  ymax = -dmdtda+err_dmdtda, 
                  colour = RGIId, linetype = "obs", fill = RGIId), 
              size = 0.2, alpha = 0.2) + 
  geom_line(data = melt_a |> 
              dplyr::filter(RGIId %in% glaciers[6:9]), 
            aes(date, Melt/1000, colour = RGIId, linetype = "sim")) + 
  geom_line(data = melt_obs_a, aes(start, -dmdtda, colour = RGIId, 
                                   linetype = "obs")) + 
  labs(x = "Date", y = "Melt [m weq/a]") + 
  scale_linetype_manual(name = "Source", 
                        values = c("sim" = 1, "obs" = 2)) + 
  scale_colour_viridis_d() + 
  scale_fill_viridis_d() + 
  ggtitle(paste0("MF: ", MF, "Tth: ", threshold_temperature)) + 
  theme_bw()
```

With the synthetic temperatures we cannot obtain the observed steep increase of the glacier melt. In theory, the temperature index model can be calibrated with the specific glacier volume change provided by @hugonnet_accelerated_2021 (see \\ref{fig:glaciermeltcomparison}).

### Glacier mass balance

```{r glaciermassbalance, message=FALSE, warning=FALSE}
#| fig.cap = "Glacier development and discharge." 
#| out.width = 90%
glacier_balance <- glacierBalance(melt_a_eb = melt_a_eb, 
                                  rgi_elbands = rgi_elbands)

glacier_balance <- glacier_balance |> 
  mutate(.lb = ifelse(Variable == "A_km2", 
                      ifelse(Value*(1-error_stats$sA)>0, 
                             Value*(1-error_stats$sA), 0), 
                      ifelse(Variable == "V_km3", 
                             ifelse(Value*(1-error_stats$sV)>0, 
                                    Value*(1-error_stats$sV), 0), 
                             ifelse(Variable == "Q_m3a", 
                                    ifelse(Value*(1-error_stats$sQglmelt)>0,
                                           Value*(1-error_stats$sQglmelt), 0), 
                                    ifelse(Variable == "Qimb_m3a", 
                                           Value*(1-error_stats$sImbal), NA)))), 
         .ub = ifelse(Variable == "A_km2", 
                      Value*(1+error_stats$sA), 
                      ifelse(Variable == "V_km3", 
                             Value*(1+error_stats$sV), 
                             ifelse(Variable == "Q_m3a", 
                                    Value*(1+error_stats$sQglmelt), 
                                    ifelse(Variable == "Qimb_m3a", 
                                           Value*(1+error_stats$sImbal), NA)))))

ggplot(glacier_balance |> 
         dplyr::filter(RGIId %in% glaciers[6:9], 
                       Hyear > min(Hyear) & Hyear < max(Hyear), 
                       Variable %in% c("A_km2", "V_km3", "Q_m3a", "Qimb_m3a"))) + 
  geom_ribbon(aes(Hyear, ymin = .lb, ymax = .ub, fill = RGIId), 
              alpha = 0.2, colour = NA) + 
  geom_line(aes(Hyear, Value, colour = RGIId)) + 
  facet_wrap("Variable", scales = "free_y") + 
  scale_colour_viridis_d() + 
  scale_fill_viridis_d() + 
  theme_bw()
```

We now have glacier discharge (Q_m3s) and the unsustainable contribution to glacier discharge, the imbalance ablation (Qimb_m3a) which is negative for glacier loss and positive for growing glaciers. We are only interested in the contribution of imbalance ablation to river discharge, that is, only the negative part of Qimb_m3a is relevant to us.

# Next steps

Now we have glacier discharge and glacier imbalance ablation. How do we include these in our modelling workflow? [Glacier vignette 4](glaciers-04-glaciers-RSM.html) explains how glaciers are treated in the hydraulic/hydrologic model RS Minerve and details a workflow to implement glacier discharge in a model in RS Minerve.

```{r, include=FALSE, eval=FALSE}
# Saving variables for continuing demo in vignette 4
melt_d <- melt |> 
  pivot_longer(-date, names_to = "ID", values_to = "melt") |> 
  separate(ID, into = c("RGIId", "elB"), sep = "_") |> 
  group_by(date, RGIId) |> 
  summarize(melt = sum(melt)) |> 
  ungroup()
saveRDS(melt_d |> rename(M_mmd = melt), file = "vignettes/daily_glacier_melt.rds")

saveRDS(glacier_balance, file = "vignettes/glacier_balance.rds")
```

# Other glacier modeling resources worth mentioning

Of course there are many more glacier utilities out there which are worth checking out if you are interested in more in-depth glacier modeling.

## Open Global Glacier Model (OGGM)

The [Open Global Glacier Model](https://oggm.org/){target="_blank"} is a community project which provides glacier modelling tools in python and educational resources. The modeling framework includes comprehensive tutorials and data-pre and post-processing tools.

# References
